---
title: "COMPSCIX 415.2 Homework 3"
author: "Rajat Jain"
date: "June 25th, 2018"
output:
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
# For NYC Flights data.
library(nycflights13)
# For display grids.
library(grid)
library(gridExtra)
```

## Section 3.8.1 Exercises

1. **What is the problem with this plot? How could you improve it?**
```{r 3_8_1_1_q, fig.height=4}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

\pagebreak
This plot has many overlapping point which do not show the areas of concentration properly. It can be inproved by adding jitter to the position adjustment.
```{r 3_8_1_1, fig.height=4}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```


2. **What parameters to `geom_jitter()` control the amount of jittering?**

The parameters to `geom_jitter()` which control the amount of jittering are `width` and `height`.
\newline

3. **Compare and contrast `geom_jitter()` with `geom_count()`.**

`geom_jitter` randomly moves the overlapping points slightly to avoid overlapping whereas, `geom_count` counts the overlapping points at a given point and maps them to the size of a single point. The determinism of `geom_count` makes it useful in discrete situations, but it does not work when the points are not exactly overlapping but are very close.
\newline

4. **What’s the default position adjustment for `geom_boxplot()`? Create a visualisation of the `mpg` dataset that demonstrates it.**

The default position adjustment for `geom_boxplot()` is `position_dodge`. Here is a demonstration:
```{r 3_8_1_4, fig.height=7}
my_plot <- ggplot(data = mpg, mapping = aes(x = fl, y = hwy, fill = drv))

# Plot with default position adjustment.
dflt <- my_plot + ggtitle("Position: Default") +
  geom_boxplot()

# Plot with dodge position adjustment.
ddge <- my_plot + ggtitle("Position: Dodge") + 
  geom_boxplot(position = "dodge")

# Plot with identity position adjustment.
idnt <- my_plot + ggtitle("Position: Identity") + 
  geom_boxplot(position = "identity")

# Plot with jitter position adjustment.
jter <- my_plot + ggtitle("Position: Jitter") +
  geom_boxplot(position = "jitter")

grid.arrange(dflt, ddge, idnt, jter, nrow = 2, ncol = 2)
```

In the above plots position adjustment `Dodge` generated the plot looking exactly as with default position adjustment.
\pagebreak

## Section 3.9.1 Exercises

2. **What does `labs()` do? Read the documentation.**

`labs()` function from `ggplot2` package is used to modify axis, legend, and plot labels.
\newline

4. **What does the plot below tell you about the relationship between city and highway mpg? Why is `coord_fixed()` important? What does `geom_abline()` do?**
```{r 3_9_1_4_q}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

This plot tells us that the city and highway mpg are positively correlated, meaning if a vehicle has a higher city mpg, it is also expected to have a higher highway mpg.

`coord_fixed()` is important because both city and highway mpg have the same unit and hence one unit of each should be represented by the same length on the axes for the best visualization. `coord_fixed()` does exactly that for us.

`geom_abline()` creates a reference line shown on the plot. Since it is called without `intercept` and `slope` arguments, it generates a default reference line passing through the origin and with a slope of 1 (45&deg;). So this line represents all the point on the plot where city mpg would be equal to highway mpg. From the plot we can also observe that all the plotted points are above this line, which means that the highway mpg is always higer than the city mpg. If we observe closely, we can also see that the scatter plot is more or less parallel to the reference line. This tells us that the highway mpg is higher than the city mpg for a constant additive offset.
\newline



## Section  4.4 Exercises

1. **Why does this code not work?**
```{r 4_4_1_q, eval=FALSE}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found```
```
**Look carefully! (This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)**

This code does not work because the name of the variable while printing it is mis-spelled.
```{r 4_4_1}
my_variable <- 10
my_variable
```
The above code is its fixed version which works.
\newline

2. **Tweak each of the following R commands so that they run correctly:**
```{r 4_4_2_q, eval=FALSE}
library(tidyverse)

ggplot(dota = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8)
filter(diamond, carat > 3)
```

Here is the fixed and working code:
```{r 4_4_2, fig.height=3.5}
# Package tidyverse is pre-loaded in this document.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)

filter(diamonds, carat > 3)
```
\pagebreak


## Section 5.2.4 Exercises

1. **Find all flights that**

    1. Had an arrival delay of two or more hours
```{r 5_2_4_1_1}
filter(flights, arr_delay >= 120)
```
&nbsp;

    2. Flew to Houston (IAH or HOU)
```{r 5_2_4_1_2}
filter(flights, dest %in% c("IAH", "HOU"))
```
&nbsp;

    3. Were operated by United, American, or Delta
```{r 5_2_4_1_3}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```
&nbsp;

    4. Departed in summer (July, August, and September)
```{r 5_2_4_1_4}
filter(flights, month %in% 7:9)
```
&nbsp;

    5. Arrived more than two hours late, but didn’t leave late
```{r 5_2_4_1_5}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```
&nbsp;

    6. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r 5_2_4_1_6}
filter(flights, dep_delay >= 60 & dep_delay - arr_delay > 30)
```
&nbsp;

    7. Departed between midnight and 6am (inclusive)
```{r 5_2_4_1_7}
filter(flights, dep_time <= 600 | dep_time == 2400)
```
\pagebreak

3. **How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?**

```{r 5_2_4_3}
summary(flights)
```

`r sum(is.na(flights$dep_time))` flights have missing `dep_time`. Other variables with missing values are: `dep_delay`, `arr_time`, `arr_delay` & `air_time`.

Since only those variables seem to have missing values which are related to actual flight instances and not just the scheduled details, they may represent the cancelled flights. Alternatively, they could just be errors in data entry while recording those values at flight departure and/or arrival. Or a mix of both.
\pagebreak


4. **Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)**

```{r 5_2_4_4}
NA ^ 0

NA | TRUE

FALSE & NA

NA * 0
```

`NA ^ 0` is not missing because anything to the power of 0 is 1 irrespective of what the value is. However, this is not true for `Inf ^ 0`. It is still indeterminate form. So this seems to be an incorrect evaluation by R.

`NA | TRUE` is not missing because the `|` expression evaluates to `TRUE` if either parts of the expression evaluate to `TRUE` and since `TRUE` is always `TRUE` it doesn't matter what the other part is.

Similarly, `FALSE & NA` is not missing because `&` expression evaluates to `FALSE` if either parts of the expression evaluates to `FALSE` irrespective of the value of the other part.

`NA * 0` evaluates to `NA` because NA could take any value including `Inf` and `Inf * 0` is indeterminate form.
\pagebreak


## Section 5.4.1 Exercises

1. **Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.**

```{r 5_4_1_1}
# Define column names vector.
col_names <- c("dep_time",  "dep_delay", "arr_time", "arr_delay")


#1. Using [] with column indexes.
flights[, c(4, 6, 7, 9)]


#2. Using [] with column names vector.
flights[, col_names]


#3. Using select with column indexes.
select(flights, 4, 6, 7, 9)


#4. Using select with column names vector.
select(flights, col_names)


#5. Using select with column names mentioned as ... arguments.
select(flights, dep_time,  dep_delay, arr_time, arr_delay)


#6. Using select with pipe operator and column indexes.
flights %>% select(4, 6, 7, 9)


#7. Using select with pipe operator and column names vector.
flights %>% select(col_names)


#8. Using select with pipe operator and column names mentioned as ... arguments
flights %>% select(dep_time,  dep_delay, arr_time, arr_delay)


#9. Using starts_with() select helper.
flights %>% select(starts_with("dep_"), starts_with("arr_"))


#10. Using one_of() select helper.
flights %>% select(one_of(col_names))
```
&nbsp;


3. **What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?**
```{r 5_4_1_3_q, fig.height=3.5}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```

`one_of()` is a Select Helper function from `dplyr` package. It allows for guessing or subset-matching. (Courtesy: [StackOverflow](https://stackoverflow.com/questions/45865892/why-is-one-of-called-that)). It returns all the columns with names which match the vector provided to it. For example, in conjunction with the vector above, it could be used with `select()` as below:

```{r 5_4_1_3}
flights %>% select(one_of(vars))
```