---
title: "COMPSCIX 415.2 Homework 1"
author: "Rajat Jain"
date: "6/5/2018"
output:
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## My Repository

My Github repository for my assignments can be found at this URL: [https://github.com/rajatmnnit/compscix-415-2-assignments](https://github.com/rajatmnnit/compscix-415-2-assignments)


## Libraries

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```


## Exploring Data

### Load data from Package

```{r data}
data("WorldCities")
```

### Data Summary - Question# 1
WorldCities data set contains `r nrow(WorldCities)` observations and `r ncol(WorldCities)` variables. Some of the variable names are: `r colnames(WorldCities)`


### A quick glimpse at the data
```{r datapreview}
# Using glimpse function from dplyr
glimpse(WorldCities)
```


## Extraction

### Top 200 Rows
```{r filter_worldcities}
WorldCities <- head(WorldCities, 200) # 200 rows
```

### Countries
```{r countries}
country_col <- WorldCities$country
unique(country_col)
```

### Regions - Question# 2
```{r regions}
unique(WorldCities$region)
```

### The tidy way - Question# 3
```{r unique_countries}
WorldCities %>% select(country) %>% unique()
```


## Visualize

```{r pop_dist}
WorldCities %>% ggplot(aes(x = population)) +
  geom_histogram()
```

### Population Distribution - Question# 4
```{r pop_dist_titled}
WorldCities %>% ggplot(aes(x = population)) +
  geom_histogram() +
  xlab("City Population") +
  ylab("#Cities") + 
  ggtitle("Distribution of Population in Cities") + 
  theme_bw()
```